Three objections to a novel paradigm in social media effects research

### 2022-11-10 ms.Rmd
Fixed abstract

---
title: Three objections to a novel paradigm in social media effects research
shorttitle: Three objections
leftheader: Three objections
author: 
  - name: Matti Vuorre
    affiliation: 1
    corresponding: yes    # Define only one corresponding author
    address: Oxford Internet Institute, University of Oxford
    email: matti.vuorre@oii.ox.ac.uk
    role:
      - Conceptualization
      - Formal analysis
      - Software
      - Visualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name: Niklas Johannes
    affiliation: 1
    address: Oxford Internet Institute, University of Oxford
    email: niklas.johannes@oii.ox.ac.uk
    role:
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name: Andrew K. Przybylski
    affiliation: 1
    address: Oxford Internet Institute, University of Oxford
    email: andy.przybylski@oii.ox.ac.uk
    role:
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
affiliation:
  - id: 1
    institution: Oxford Internet Institute, University of Oxford
authornote: |
  A non-peer reviewed pre-print of this work is available at <https://psyarxiv.com/dpuya>.
  All code supporting this manuscript is available at <https://doi.org/10.5281/zenodo.6838890>.
abstract: |
  Here, we outline three objections to that research agenda. On a methodological level, the key empirical results of this programme---proportions of the population of individuals with negative, null, and positive social media effects---are not appropriately estimated and reported. On a theoretical level, these results do little to advance our understanding of social media and its psychological implications. On a  paradigmatic level, this "personalized media effects paradigm" [@valkenburgSocialMediaUse2021, p.74] cannot inform inferences about individuals and therefore does not deliver what it claims. In this work we express our concern that this research approach may be contributing to confusing messaging to both societal stakeholders and scientists investigating how social media and well-being might be related. It is our sincere hope that describing these objections directly will prompt the field to work together in adopting better practices to ultimately develop a better understanding of well-being in the digital age. The study of social media effects on psychological well-being has reached an impasse: Popular commentators confidently assert that social media are bad for users but research results are mixed and have had little practical impact. In response, one research group has proposed a path forward for the field that moves beyond studying population averages to find effects that are specific to individuals.
keywords: social media, media effects, well-being, variation, paradigm
wordcount: "2,802"
bibliography: person-specific-paradigm.bib
floatsintext: yes
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass: apa7
classoption: jou
output: 
  papaja::apa6_pdf:
    number_sections: true
  papaja::apa6_docx: 
    number_sections: true
editor_options: 
  chunk_output_type: console
---

>>>>

---
title: Three objections to a novel paradigm in social media effects research
shorttitle: Three objections
leftheader: Three objections
author: 
  - name: Matti Vuorre
    affiliation: 1
    corresponding: yes    # Define only one corresponding author
    address: Oxford Internet Institute, University of Oxford
    email: matti.vuorre@oii.ox.ac.uk
    role:
      - Conceptualization
      - Formal analysis
      - Software
      - Visualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name: Niklas Johannes
    affiliation: 1
    address: Oxford Internet Institute, University of Oxford
    email: niklas.johannes@oii.ox.ac.uk
    role:
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name: Andrew K. Przybylski
    affiliation: 1
    address: Oxford Internet Institute, University of Oxford
    email: andy.przybylski@oii.ox.ac.uk
    role:
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
affiliation:
  - id: 1
    institution: Oxford Internet Institute, University of Oxford
authornote: |
  A non-peer reviewed pre-print of this work is available at <https://psyarxiv.com/dpuya>.
  All code supporting this manuscript is available at <https://doi.org/10.5281/zenodo.6838890>.
abstract: |
  The study of social media effects on psychological well-being has reached an impasse: Popular commentators confidently assert that social media are bad for users but research results are mixed and have had little practical impact. In response, one research group has proposed a path forward for the field that moves beyond studying population averages to find effects that are specific to individuals. Here, we outline three objections to that research agenda. On a methodological level, the key empirical results of this programme---proportions of the population of individuals with negative, null, and positive social media effects---are not appropriately estimated and reported. On a theoretical level, these results do little to advance our understanding of social media and its psychological implications. On a  paradigmatic level, this "personalized media effects paradigm" [@valkenburgSocialMediaUse2021, p.74] cannot inform inferences about individuals and therefore does not deliver what it claims. In this work we express our concern that this research approach may be contributing to confusing messaging to both societal stakeholders and scientists investigating how social media and well-being might be related. It is our sincere hope that describing these objections directly will prompt the field to work together in adopting better practices to ultimately develop a better understanding of well-being in the digital age. 
keywords: social media, media effects, well-being, variation, paradigm
wordcount: "2,802"
bibliography: person-specific-paradigm.bib
floatsintext: yes
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass: apa7
classoption: jou
output: 
  papaja::apa6_pdf:
    number_sections: true
  papaja::apa6_docx: 
    number_sections: true
editor_options: 
  chunk_output_type: console
---
